---
title: 'Project 2: Modeling, Testing, and Predicting'
author: "SDS348"
date: "5/7/21"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})

knitr::opts_chunk$set(echo = TRUE, eval = TRUE,fig.align="center",warning=FALSE,message=FALSE,fig.width=8, fig.height=5, linewidth=60)
options(tibble.width = 100,width = 100)
library(tidyverse)
```

# Modeling

## Guidelines and Rubric

- **0. (5 pts)** Introduce your dataset and each of your variables (or just your main variables if you have lots) in a paragraph. What are they measuring? How many observations?
```{r}
library(tidyverse)
library(AER)
phd <- data("PhDPublications")
summary(PhDPublications)

```
The data I selected is PhD publications. It has six variables and 915 observations. The variables are number of articles published during last 3 years of PhD (articles), gender, kids less than six years old (kids), married, prestige of the program (prestige) and number of articles published by students mentor (mentor).

- **1. (15 pts)** Perform a MANOVA testing whether any of your numeric variables (or a subset of them, if including them all is unreasonable or doesn't make sense) show a mean difference across levels of one of your categorical variables (3). If they do, perform univariate ANOVAs to find response(s) showing a mean difference across groups (3), and perform post-hoc t tests to find which groups differ (3). Discuss the number of tests you have performed, calculate the probability of at least one type I error (if unadjusted), and adjust the significance level accordingly (bonferroni correction) before discussing significant differences (3). Briefly discuss some of the MANOVA assumptions and whether or not they are likely to have been met here (no need for anything too in-depth) (2).
```{r}

man1 <- manova(cbind(articles,kids,mentor)~prestige,data = PhDPublications)

summary(man1)

summary.aov(man1)

#pairwise.t.test(PhDPublications$kids,PhDPublications$prestige, p.adj = "none")
#pairwise.t.test(kids,prestige, p.adj = "none")
#pairwise.t.test(mentor,prestige, p.adj = "none")
```
A manova test was performed to test if any of my numerical variables would show a mean different among levels on one of my categorical variables. The pvalue is 4.19e-14 which shows there is a mean different. I muted the post.t test because it didn't print anything useful and it was long.


- **3. (40 pts)** Build a linear regression model predicting one of your response variables from at least 2 other variables, including their interaction. Mean-center any numeric variables involved in the interaction.

    - Interpret the coefficient estimates (do not discuss significance) (10)
    - Plot the regression using `ggplot()` using geom_smooth(method="lm"). If your interaction is numeric by numeric, refer to code in the slides to make the plot or check out the `interactions` package, which makes this easier. If you have 3 or more predictors, just chose two of them to plot for convenience. (10)
    - What proportion of the variation in the outcome does your model explain? (4)
    - Check assumptions of linearity, normality, and homoskedasticity either graphically or using a hypothesis test (5)
    - Regardless, recompute regression results with robust standard errors via `coeftest(..., vcov=vcovHC(...))`. Discuss significance of results, including any changes from before/after robust SEs if applicable. (10)
```{r}
library(tidyverse)
fit <- lm(prestige~articles+mentor,data = PhDPublications)
summary(fit)
coef(fit)

fit %>% ggplot(aes(prestige,articles+mentor))+geom_point()+geom_smooth(method = "lm",se = F)
coeftest(fit,vcov=vcovHC(fit))
```
  
Our model line  has a positive slope but we appear to have outliers. Graphically, it does not seem linearity, normality, and homoskedasticity are met. Our intercept means that on average univeristies have a prestige of 2.870253 while a PhD student has -0.003558  publications in the past three years while mentors have published 0.027247 publications. The model has a propotion of 0.06786 of the variation in the outcome. The SE seems to decrease. 


- **5. (30 pts)** Fit a logistic regression model predicting a binary variable (if you don't have one, make/get one) from at least two explanatory variables (interaction not necessary). 

    - Interpret coefficient estimates in context (10)
    - Report a confusion matrix for your logistic regression (5)
    - Compute and discuss the Accuracy, Sensitivity (TPR), Specificity (TNR), Precision (PPV), and AUC of your model (5)
    - Using ggplot, make a density plot of the log-odds (logit) colored/grouped by your binary outcome variable (5)
    - Generate an ROC curve (plot) and calculate AUC (either manually or with a package); interpret (5)
```{r}
class_diag <- function(probs,truth){ 
  #CONFUSION MATRIX: CALCULATE ACCURACY, TPR, TNR, PPV 
  if(is.character(truth)==TRUE) truth<-as.factor(truth) 
  if(is.numeric(truth)==FALSE & is.logical(truth)==FALSE) truth<-as.numeric(truth)-1 
  tab<-table(factor(probs>.5,levels=c("FALSE","TRUE")),factor(truth, levels=c(0,1))) 
  acc=sum(diag(tab))/sum(tab) 
  sens=tab[2,2]/colSums(tab)[2] 
  spec=tab[1,1]/colSums(tab)[1] 
  ppv=tab[2,2]/rowSums(tab)[2] 
  
#CALCULATE EXACT AUC 
  ord<-order(probs, decreasing=TRUE) 
  probs <- probs[ord]; truth <- truth[ord] 
  TPR=cumsum(truth)/max(1,sum(truth))  
  FPR=cumsum(!truth)/max(1,sum(!truth)) 
  dup <-c(probs[-1]>=probs[-length(probs)], FALSE) 
  TPR <-c(0,TPR[!dup],1); FPR<-c(0,FPR[!dup],1) 
  n <- length(TPR) 
  auc <- sum( ((TPR[-1]+TPR[-n])/2) * (FPR[-1]-FPR[-n])) 
  data.frame(acc,sens,spec,ppv,auc) 
}

```
    

- **6. (25 pts)** Perform a logistic regression predicting the same binary response variable from *ALL* of the rest of your variables (the more, the better!) 

    - Fit model, compute in-sample classification diagnostics (Accuracy, Sensitivity, Specificity, Precision, AUC), and interpret (5)
    - Perform 10-fold (or repeated random sub-sampling) CV with the same model and report average out-of-sample classification diagnostics (Accuracy, Sensitivity, Specificity, Precision, and AUC); interpret AUC and compare with the in-sample metrics (10)
    - Perform LASSO on the same model/variables. Choose lambda to give the simplest model whose accuracy is near that of the best (i.e., `lambda.1se`). Discuss which variables are retained. (5)
    - Perform 10-fold CV using only the variables lasso selected: compare model's out-of-sample AUC to that of your logistic regressions above (5)

