---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r eval=F}
## paste this chunk into the ```{r setup} chunk at the top of your project 1 .Rmd file

knitr::opts_chunk$set(echo = TRUE, eval = TRUE, fig.align = "center", warning = F, message = F,
tidy=TRUE, tidy.opts=list(width.cutoff=60), R.options=list(max.print=100))
```

## Project one SDS 348
## Ali Chaghani
I decided on two data sets from carData and fivethirtyeight to see if states with 
a higher percentage of high school graduates who took a SAT have higher insurance 
premiums. Overall, the variables my joint dataframe have state, Number of drivers involved in fatal collisions per billion miles, insurance_premiums, population, SAT verbal scores, and percent of high school graduates who took the SAT.
I choose this as my project because I am currently teaching my bad driver best friend how to drive. 

```{r}
library(fivethirtyeight)
library(carData)
library(tidyverse)
library(cluster)

df <- subset(bad_drivers, select = -c(perc_no_previous))
df_2 <- cbind( state = df$state,States)
df_j <- left_join(df,df_2)
new_df_j <- subset(df_j,select = c(state,num_drivers,insurance_premiums,pop,SATV,percent))
head(new_df_j)
```
These states are SAT scores above the mean SAT score of the USA. 
```{r}
#SD
new_df_j %>% group_by(percent,insurance_premiums) %>% summarise_if(is.numeric,list(sd=sd), na.rm = T)

#Variance
new_df_j %>% group_by(percent,insurance_premiums) %>% summarise_if(is.numeric,list(var=var), na.rm = T)

#quantile
new_df_j %>% summarise_if(is.numeric,list(Q1=quantile),probs = .25,na.rm=T)
new_df_j %>% summarise_if(is.numeric,list(Q2=quantile),probs = .5,na.rm=T)
new_df_j %>% summarise_if(is.numeric,list(Q3=quantile),probs = .75,na.rm=T)
new_df_j %>% summarise_if(is.numeric,list(Q4=quantile),probs = 1,na.rm=T)

#min and max 
min_df <- new_df_j %>% summarise_if(is.numeric,list(min=min),na.rm = T)
min_df %>% pivot_longer(cols = 2:5,names_to = "stat",values_to = "Value")

max_df <- new_df_j %>% summarise_if(is.numeric,list(max=max),na.rm = T)
max_df %>% pivot_longer(cols = 2:5,names_to = "stat",values_to = "Value")

new_df_j %>% group_by(percent,insurance_premiums)%>%summarize_if(is.numeric,list(mean=mean),na.rm=T)
```
I found the basic five stats for my data. 

```{r}
ggplot(data=new_df_j, mapping = aes(x = percent, y = insurance_premiums)) + 
  geom_point(aes(color = state)) +
  theme_bw()

ggplot(data = new_df_j,mapping = aes(x = percent, y = insurance_premiums)) +
  geom_line()

```
As you can see from the graphs there is no clear relationship between graduating high school and taking the SAT to predict your insurance premium. 

```{r}
pam1<-new_df_j%>%pam(k=3)
pam1$silinfo$avg.width

pam_dat<-new_df_j%>%select(insurance_premiums,percent,num_drivers)
sil_width<-vector()
for(i in 2:10){  
  pam_fit <- pam(pam_dat, k = i)  
  sil_width[i] <- pam_fit$silinfo$avg.width  
}
ggplot()+geom_line(aes(x=1:10,y=sil_width))+scale_x_continuous(name="k",breaks=1:10)
```
Using two would be the best cluster. 

Started this very late and this isn't my best work. Will do better on the next one. 
